---
title: "Getting Started With `sommer`"
author: "Peter von Rohr"
date: "2018-08-06"
output: html_notebook
---

## Disclaimer
This notebook gives a few scratch ideas on how to use the package `sommer` for a simple linear multivariate mixed model analysis. We want to see different effects of a multivariate versus a univariate analysis.


## Installation
The first step is to install the package that we want to experiment with

```{r}
if (!require(sommer))
  install.packages("sommer")
```


## Data Simulation
The tests on how to use the package `sommer` and possibly some other helper packages is tested by a simple analysis of a small simulated dataset. This section describes how the data is simulated.

### Parameters
We start by fixing some parameters

```{r, echo=FALSE}
set.seed(5098)
vp <- 36
vg <- 9
ve <- vp - vg
h2 <- vg / vp
```

The following table gives an overview over the assumed parameter values

```{r, echo=FALSE}
df_par <- dplyr::tibble(Parameter = c("phenotypic variance",
                                      "genetic additive variance",
                                      "residual variance",
                                      "heritability"),
                        Value = c(vp,
                                  vg,
                                  ve,
                                  h2))
knitr::kable(df_par, align = c('l','r'))
```


### Pedigree
```{r}
nsire <- 3
ndam <- 9
nfounder <- nsire + ndam
```

The pedigree consists of `r nsire` unrelated sires which are mated to `r ndam` different dams. In total, we get `r nfounder` so-called founder animals. Those are all animals without known parents. Based on these animals, we can create the founder pedigree

```{r}
suppressPackageStartupMessages( require(pedigreemm) )
ped_founder <- pedigree(sire = rep(NA, nfounder), dam = rep(NA, nfounder), label = as.character(1:nfounder))
ped_founder
```

Founder animals are mated to produce the first offspring generation. The number of offspring for each parent is determined in a rather arbitrary way. Together with the founders this generates the pedigree up until the first generation.

```{r}
mat_mating_firstgen <- matrix(NA, nrow = nfounder, ncol = nfounder)
mat_mating_firstgen[1,4] <- 2
mat_mating_firstgen[1,5] <- 1
mat_mating_firstgen[1,6] <- 1
mat_mating_firstgen[1,7] <- 1
mat_mating_firstgen[2,8] <- 2

vec_nr_offspring_sire <- c(4,3,2)

# ped_firstgen <- pedigree(sire = c(rep(1,)))
```


```{r, echo=FALSE, results='asis'}
cat(rmddochelper::get_latest_change())
```

